---
title: "Group 5 Report"
author: "Dagfinn, Mathilde, Joanna, Ingvild"
date: "2023-09-15"
output: html_document
---

Load library
```{r, echo=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(patchwork)
library(skimr)
library(naniar)
library(GGally)
```

Read file
```{r, echo=FALSE, message=FALSE}
OurData <- read_delim(here("Data", "tidy_exam_data.txt"))
```

## Description of data

### Tidying data

## Wrangling data

## Plotting data
### Exploring correlated measurements
```{r, echo=FALSE, warning=FALSE, message=FALSE}
multiple_correlations_plot <- ggpairs(OurData[c("BMI", 
                    "Age", 
                    "Education", 
                    "GA.at.outcome", 
                    "Birthweight",
                    "IL6_baseline",
                    "IL8_baseline")])

multiple_correlations_plot
```


## Analysis
Lastly, we analyzed our data focusing on birthweight and birth outcome. First we checked whether birth outcome was dependent of the enrollment center. It was not (p=XXX) 

```{r}
knitr::opts_chunk$set(echo = F,
                      warning = F)
Enroll.Center_birthweight <- ggplot(OurData,
                                aes(x = Enroll.Center, y = Birthweight, fill = Enroll.Center)) +
  geom_boxplot(aes(color = Enroll.Center)) +
  theme_minimal() +
  labs(title = "Enroll.Center vs. birthweight")
Enroll.Center_birthweight

OurData %>%
  aov(Birthweight~Enroll.Center, data = .) %>%
  broom::tidy()
```

Second, birth outcome was neither dependent of BMI of the patient (p=).
```{r}
Hist_BMI_Birth.outcome <- ggplot(OurData, aes(x=BMI)) +
  geom_histogram(aes(fill=Birth.outcome), bins=30, alpha=0.7, position="identity") + 
  facet_wrap(~ Birth.outcome, scales="free_y") +
  labs(title="Distribution of BMI by Birth Outcome", 
       x="BMI", y="Count") +
  theme_minimal()

print(Hist_BMI_Birth.outcome)

OurData %>% 
  kruskal.test(BMI~Birth.outcome, data = .) %>%
  broom::tidy()
```

Third, 

##Session Info
```{r}
sessionInfo()
```



